---
title: "CedricScherer_TutorialBeautifulGGplotting"
author: "Dominik Ziaja"
date: "2022-10-26"
output: html_document
---


Markdown file focusing on the tutorial of cedric scherer (beautiful work!) for
plotting well visualised plots using ggplot. Probably rather snippets and part
I find interesting/didn't know yet than a thorough completion of the whole tutorial.

Tutorial can be found here: https://www.cedricscherer.com/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/

# Packages recommended to use:

* *ggplot2*
* *tidyverse*
  * *dplyr* - data wrangling
  * *tibble* - modern data frames
  * *tidyr* - data cleaning
  * *forcats* - for handlings factors
* *colorspace* - manipulating colors
* *corrr* - calculating correlation matrices
* *cowplot* - composing ggplots
* *ggdark* - themes and inverting colors
* *ggforce* - sina plots and some other cool stuff
* *ggrepel* - text labeling
* *ggridges* - ridge plots (waterfall plots)
* *ggsci* - color palettes
* *ggtext* - text rendering
* *ggthemes* - additional themes
* *grid* - creating graphical objects
* *gridExtra* - additional grid graphics functions
* *patchwork* - multi-panel plots - Where exactly difference to cowplot?
* *rcartcolor* - color palettes
* *scico* - perceptional uniform palettes
* *showtext* - custom fonts
* *shiny* - interactive apps
* list of packages for interactive visualisations:
  * *charter*
  * *echarts4r*
  * *ggiraph*
  * *highcharter*
  * *plotly*


# Loading in the packages and dataset

```{r}
library(tidyverse)
library(showtext)
```


```{r}
chic <- read_csv("https://raw.githubusercontent.com/z3tt/ggplot-courses/main/data/chicago-nmmaps-custom.csv")
```

```{r}
chic %>%
  glimpse()
```
# Set the default theme and add fonts

```{r}

font_add_google("Playfair Display",
                "Playfair")
font_add_google("Bangers", "Bangers")
font_add_google("Roboto Condensed", "Roboto Condensed")

theme_set(theme_bw(base_size = 12, base_family = "Roboto Condensed")) # set default font for all plots
showtext_auto()
```

# Modify cartesian axis

We can force graphs to start at the origin by using expand limits
```{r}
chic %>%
  filter(temp > 25, o3 > 20) %>%
  ggplot(data = ., aes(x = temp, y = o3)) +
  geom_point(color = "darkcyan") +
  labs(x = "Temperature higher than 25°F",
       y = "Ozone higher than 20 ppb") +
  expand_limits( x = 0, y = 0)
```

Or if we want it to literally start at 0, we need to use the expand = False command

```{r}
chic %>%
  filter(temp > 25, o3 > 20) %>%
  ggplot(data = ., aes(x = temp, y = o3)) +
  geom_point(color = "darkcyan") +
  labs(x = "Temperature higher than 25°F",
       y = "Ozone higher than 20 ppb") +
  expand_limits( x = 0, y = 0) +
  coord_cartesian(expand = FALSE, clip = "off")
```

coord_fixed allows us to define the ratios between x and y axis. How many units
of x should be in one unit of y lengthwise?

```{r}
ggplot(chic, aes(x = temp, y = temp + rnorm(nrow(chic), sd = 20))) +
  geom_point(color = "sienna") +
  labs(x = "Temperature (°F)", y = "Temperature (°F) + random noise") +
  xlim(c(0, 100)) + ylim(c(0, 150)) +
  coord_fixed(ratio = 1) # one unit on x axis is the same length as one unit on the y axis
```

```{r}
ggplot(chic, aes(x = temp, y = temp + rnorm(nrow(chic), sd = 20))) +
  geom_point(color = "sienna") +
  labs(x = "Temperature (°F)", y = "Temperature (°F) + random noise") +
  xlim(c(0, 100)) + ylim(c(0, 150)) +
  coord_fixed(ratio = 1/10) # 1 unit on the x axis is as long as 10 units on the y axis
```

# Changing up the titles

```{r}
ggplot(chic, aes(x = date, y = temp)) +
  geom_point(color = "firebrick") +
  scale_y_continuous(label = function(x) {return(paste(x, "Degrees Fahrenheit"))}) + # nice trick to insert redundant text into plot
  labs(x = "Year", y = NULL,
       title = "Temperatures in Chicago between 1997 and 2001 in Degrees",
       caption = "Data: NMAPS") +
  theme(plot.title = element_text(size = 14, lineheight = .8, # lineheight changes the spacing between lines
                                  face = "bold.italic", family = "Bangers"),
        plot.caption = element_text(hjust = 0, family = "Playfair"),
        plot.title.position = "plot", # position = "plot" adds title leftoriented above the y axis and is the preferred solution
        plot.caption.position = "plot")
```
# Modifying a legend

Using the legend.position argument with a vector c(x <= 1, y <= 1) allows to 
place the legend into the plot. 0 is left or bottom, 1 is right or top
```{r}
ggplot(chic, aes(x = date, y = temp, color = season)) +
  geom_point() +
  labs(x = "Year", y = "Temperature (°F)", color = NULL) +
  theme(legend.position = c(.15, .15),
        legend.background = element_rect(fill = "transparent")) #+
  guides(color = guide_elegend(direction = "horizontal")) # makes the legend horizontal instead of vertical

```
 We can also change the background boxes in the legend
 
```{r}
ggplot(chic, aes(x = date, y = temp, color = season)) +
  geom_point() +
  labs(x = "Year", y = "Temperature (°F)") +
  theme(legend.key = element_rect(fill = "darkgoldenrod1"), # draws an rectangle in darkgoldenrod1 around the legend keys
        legend.title = element_text(family = "Playfair", # custom fonts for legend keys work best
                                    color = "chocolate",
                                    size = 14, face = 2)) +
  scale_color_discrete("Seasons:") +
  guides(color = guide_legend(override.aes = list(size = 6))) + # sets the size of legend keys
  geom_rug(show.legend = FALSE) # plots rug plots additionally, but the rug legend is not shown in the legend
```

# Add 